---
title: "01 Curating chromMAGMA and MAGMA Gene Identifiers"
output: html_document
---
*Robbin Nameki, Anamay Shetty et al.*
*August 2021*

# Introduction
The following workflow describes the process of curating chromMAGMA and MAGMA outputs to the gene identifiers EnsemblgeneIDs available from biomaRt

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```
## Libraries
```{r, eval = TRUE, message= FALSE, warning=FALSE}
library(biomaRt)
library(knitr)
```

## Sourcecode
```{r,eval = TRUE, message= FALSE, warning=FALSE}
source('Rscripts/Utils.R')
```

## Curating chromMAGMA and MAGMA gene identifiers 
Gene identifiers ID'd as 'ensembl_gene_id', 'external_gene_name','external_synonym','hgnc_symbol', 'entrez_gene_id', and 'uniprot_gn_symbol' from BiomaRt are converted to EnsemblGeneIDs. The gene list is also limited here to protein coding genes. 
```{r,eval = TRUE, warning=FALSE}
#Cleaning chromMAGMA Outputs. 

.ensembl <- useEnsembl_GRCh37()

x <- read_final_gene_results()

#x$GENE[x$FEATURE == 'snp_to_genes'] <- x$originalfeature[x$FEATURE == 'snp_to_genes']

genes <- unique(sort(x$GENE))

### ensembl_gene_id
i <- startsWith(x = genes, prefix = 'ENSG') & nchar(genes) == 15
df.1 <- data.frame(GENE = genes[i], ensembl_gene_id = genes[i])
genes <- genes[!i]

### external_gene_name
bm <- getBM(attributes = c('ensembl_gene_id', 'external_gene_name', 'gene_biotype'), mart = .ensembl, filters = 'chromosome_name', values = c(1:22,'X')); bm <- bm[bm$gene_biotype == 'protein_coding',]
bm <- bm[!is.na(bm$external_gene_name),]
ensembl_gene_id <- lapply(genes, function(g){
  return(bm$ensembl_gene_id[bm$external_gene_name == g])
})
i <- lapply(ensembl_gene_id, length) == 1
df.2 <- data.frame(GENE = genes[i], ensembl_gene_id = unlist(ensembl_gene_id[i]))
genes <- genes[!i]

### external_synonym
bm <- getBM(attributes = c('ensembl_gene_id', 'external_synonym', 'gene_biotype'), mart = .ensembl, filters = 'chromosome_name', values = c(1:22,'X')); bm <- bm[bm$gene_biotype == 'protein_coding',]
bm <- bm[!is.na(bm$external_synonym),]
ensembl_gene_id <- lapply(genes, function(g){
  return(bm$ensembl_gene_id[bm$external_synonym == g])
})
i <- lapply(ensembl_gene_id, length) == 1
df.3 <- data.frame(GENE = genes[i], ensembl_gene_id = unlist(ensembl_gene_id[i]))
genes <- genes[!i]

### hgnc_symbol
bm <- getBM(attributes = c('ensembl_gene_id', 'hgnc_symbol', 'gene_biotype'), mart = .ensembl, filters = 'chromosome_name', values = c(1:22,'X')); bm <- bm[bm$gene_biotype == 'protein_coding',]
bm <- bm[!is.na(bm$hgnc_symbol),]
ensembl_gene_id <- lapply(genes, function(g){
  return(bm$ensembl_gene_id[bm$hgnc_symbol == g])
})
i <- lapply(ensembl_gene_id, length) == 1
df.4 <- data.frame(GENE = genes[i], ensembl_gene_id = unlist(ensembl_gene_id[i]))
genes <- genes[!i]

### entrezgene_id
bm <- getBM(attributes = c('ensembl_gene_id', 'entrezgene_id', 'gene_biotype'), mart = .ensembl, filters = 'chromosome_name', values = c(1:22,'X')); bm <- bm[bm$gene_biotype == 'protein_coding',]
bm <- bm[!is.na(bm$entrezgene_id),]
ensembl_gene_id <- lapply(genes, function(g){
  return(bm$ensembl_gene_id[paste0('LOC',bm$entrezgene_id) == g])
})
i <- lapply(ensembl_gene_id, length) == 1
df.5 <- data.frame(GENE = genes[i], ensembl_gene_id = unlist(ensembl_gene_id[i]))
genes <- genes[!i]

### uniprot_gn_symbol
bm <- getBM(attributes = c('ensembl_gene_id', 'uniprot_gn_symbol', 'gene_biotype'), mart = .ensembl, filters = 'chromosome_name', values = c(1:22,'X')); bm <- bm[bm$gene_biotype == 'protein_coding',]
bm <- bm[!is.na(bm$uniprot_gn_symbol),]
ensembl_gene_id <- lapply(genes, function(g){
  return(bm$ensembl_gene_id[bm$uniprot_gn_symbol == g])
})
i <- lapply(ensembl_gene_id, length) == 1
df.6 <- data.frame(GENE = genes[i], ensembl_gene_id = unlist(ensembl_gene_id[i]))
genes <- genes[!i]

df <- rbind(df.1, df.2, df.3, df.4, df.5, df.6)
x <- merge(x = x, y = df, by = 'GENE', all.x = T)

bm <- getBM(attributes = c('ensembl_gene_id', 'external_gene_name', 'gene_biotype','chromosome_name','start_position','end_position'), mart = .ensembl, filters = 'chromosome_name', values = c(1:22,'X')); bm <- bm[bm$gene_biotype == 'protein_coding',]
x <- merge(x = x, y = bm, by = 'ensembl_gene_id', all.x = T)

str(x)
```

## Saving gene-level results with EnsemblGeneIds 
This RDS file will be used in further analysis
```{r,eval = TRUE}
#Output with cleaned names
#saveRDS(x = x, file = 'Data/final_gene_results_v1.08_wEnsemblGeneId.4.2.21.rds')
```

