---
title: "02 Generation of Gene-Level Results for MAGMA and chromMAGMA"
output: html_document
---
*Robbin Nameki, Anamay Shetty et al.*
*August 2021*

# Introduction
The following code is for further cleanup/tidying of data from 01, followed by a crucial step to assign the lowest P-value to each gene for the generation of gene-level results.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Libraries
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(reshape2)
```
## Sourcecode
```{r,eval = TRUE}
source('Rscripts/Utils.R')
```
## Tidying data from 01 
This chunk describes cleaning and tidying data from 01
```{r, eval = TRUE}
#Reading_Full_chromMAGMA and MAGMA table with tidied names
full_stat <- read_final_gene_results_wEnsembl()
#taking out gene synonyms on different chromosomes
full_stat <- full_stat %>%
  filter(CHR == chromosome_name)

#Selecting for only protein coding genes
full_stat_prot <- full_stat[complete.cases(full_stat),]

#Selecting Columns and tidying names of FEATURE & GWAS_DATA; to keep consistent language with manuscript.
full_stat_prot.gene_id_tidycols <- full_stat_prot %>%
  dplyr::select(ensembl_gene_id,external_gene_name,FEATURE,GWAS_TYPE,originalfeature, CHR,START,STOP,NSNPS,NPARAM,N,ZSTAT,P,chromosome_name,start_position,end_position)

full_stat_prot.gene_id_tidycols <- full_stat_prot.gene_id_tidycols %>% 
  filter(!grepl('snp_to_individual_exons', FEATURE)) %>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "snp_to_subsetted_enhancers", replacement = "chromMAGMA")%>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "snp_to_genes", replacement = "MAGMA")%>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "all_non_mucinous", replacement = "NMOC") %>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "clearcell", replacement = "CCOC") %>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "endometrioid", replacement = "EnOC") %>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "mucinous_all", replacement = "MOC")%>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "ser_lg_lmp", replacement = "LGSOC")%>%
  dplyr::mutate_if(is.character, 
                   stringr::str_replace_all, pattern = "serous_hg_extra", replacement = "HGSOC") %>%
  dplyr::rename(originalfeature_CHR = CHR) %>%
  dplyr::rename(originalfeature_START = START) %>%
  dplyr::rename(originalfeature_STOP = STOP) %>%
  dplyr::rename(gene_body_chromosome_name = chromosome_name) %>%
  dplyr::rename(gene_body_start_position = start_position) %>%
  dplyr::rename(gene_body_end_position = end_position)
head(full_stat_prot.gene_id_tidycols)
```

## Assigning most significant P-value (enhancers) to genes
Since multiple enhancers can be assigned to genes, the most significant P-value is assigned to genes to generate gene-level outputs for MAGMA and chromMAGMA. 
```{r, eval = TRUE}
full_stat_prot.gene_id_tidycols_maximum <- full_stat_prot.gene_id_tidycols %>%
  dplyr::group_by(FEATURE, GWAS_TYPE, external_gene_name) %>%
  dplyr::mutate(min_Pvalue = min(P)) %>% 
  filter(P == min_Pvalue) %>%
  #genehancer has duplicates of the same genes with different names so taking out dups
  filter(!duplicated(external_gene_name))
#ranking by the NEG_LOG10P
full_stat_prot.gene_id_tidycols_maximum_arranged <- full_stat_prot.gene_id_tidycols_maximum %>%
  dplyr::group_by(FEATURE, GWAS_TYPE) %>%
  mutate(NEG_LOG10P = -log10(min_Pvalue)) %>%
  dplyr::group_by(FEATURE, GWAS_TYPE) %>%
  dplyr::arrange(desc(NEG_LOG10P),.by_group = TRUE) 

#Cleaned table consisting of pcoding genes with most significant enhancer assigned
head(full_stat_prot.gene_id_tidycols_maximum_arranged)

## Cleaned table consisting of pcoding genes with most significant enhancer assigned
#saveRDS(x = full_stat_prot.gene_id_tidycols_maximum_arranged, file = 'Data/Gene_Level.4.9.21.rds')
```


```{r, include = FALSE}
full_stat_prot.gene_id_tidycols_maximum_arranged$originalfeature_CHR <- paste0('chr',full_stat_prot.gene_id_tidycols_maximum_arranged$originalfeature_CHR)
full_stat_prot.gene_id_tidycols_maximum_arranged$gene_body_chromosome_name <- paste0('chr',full_stat_prot.gene_id_tidycols_maximum_arranged$gene_body_chromosome_name)
  
histotype = c('HGSOC','LGSOC','CCOC','NMOC','MOC','EnOC')
for(i in histotype) {
#chromMAGMA
  assign(paste0('chromMAGMA_',i),
         full_stat_prot.gene_id_tidycols_maximum_arranged %>%
           filter(FEATURE == 'chromMAGMA') %>%
           filter(GWAS_TYPE == paste(i)))
  
#chromMAGMA_Loop
  assign(paste0('Loop_chromMAGMA_',i),
       get(paste0('chromMAGMA_',i)) %>%
      mutate(gene_body_start = gene_body_start_position -100) %>%
      mutate(gene_body_stop = gene_body_start_position + 100) %>%
    ungroup() %>%
      dplyr::select(originalfeature_CHR,
                originalfeature_START,
                originalfeature_STOP,
                gene_body_chromosome_name,
                gene_body_start,
                gene_body_stop,
                external_gene_name,
                NEG_LOG10P) 
  )


#chromMAGMA_Bed
  assign(paste0('Bed_chromMAGMA_',i),
       get(paste0('chromMAGMA_',i)) %>%
    ungroup() %>%
      dplyr::select(originalfeature_CHR,
                originalfeature_START,
                originalfeature_STOP,
                P) 
  )

  
#chromMAGMA_Loop_sig
  assign(paste0('Loop_chromMAGMA_sig',i),
       get(paste0('chromMAGMA_',i)) %>%
         group_by(FEATURE, GWAS_TYPE) %>%
         filter(P < 0.05 / n()) %>% 
      mutate(gene_body_start = gene_body_start_position -100) %>%
      mutate(gene_body_stop = gene_body_start_position + 100) %>%
         ungroup() %>%
         dplyr::select(originalfeature_CHR,
                originalfeature_START,
                originalfeature_STOP,
                gene_body_chromosome_name,
                gene_body_start,
                gene_body_stop,
                external_gene_name,
                NEG_LOG10P) 
  )

#chromMAGMA_Bed_sig
assign(paste0('Bed_chromMAGMA_sig',i),
     get(paste0('chromMAGMA_',i)) %>%
       group_by(FEATURE, GWAS_TYPE) %>%
       filter(P < 0.05 / n()) %>%
  ungroup() %>%
    dplyr::select(originalfeature_CHR,
              originalfeature_START,
              originalfeature_STOP,
              P) 
)

}
```





  
  















