---
title: "03 Linkage Disequilibrium Level Results for MAGMA and chromMAGMA"
output: html_document
---
*Robbin Nameki, Anamay Shetty et al.*
*August 2021*

# Introduction
The following workflow describes assigning linkage disequilibrium bins to chromMAGMA and MAGMA results. Linkage disequiliberium information is from https://bitbucket.org/nygcresearch/ldetect-data/src/master/. 

This data will later be used to identify unique linkage disequiliberium regions identified by chromMAGMA, in comparison to MAGMA in 04. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Libraries
```{r, message = FALSE, warning = FALSE}
library(biomaRt)
library(tidyverse)
library(valr)
```
## Sourcecode
```{r,eval = TRUE}
source('Rscripts/Utils.R')
```
## Reading files
```{r, eval = TRUE}
.ensembl <- useEnsembl_GRCh37()

gene_file <- read_gene_level()

ld_bins <- read_ld_bins()  %>%
    mutate(ID = seq(nrow(.)))
```

## Functions 
used for the assignments of MAGMA and chromMAGMA results to linkage disequiliberium bins
```{r, eval = TRUE}
# This looks at the MAGMA and chromMAGMA referenced genes ----
run_intersection <- function(df, old_col_names, suffix) {
  
  colnames(df)[colnames(df) %in% old_col_names] <- c("chrom", "start", "end")
  
  df <- df %>%
    mutate(chrom = paste0("chr", chrom)) %>%
    bed_intersect(ld_bins, suffix = c("", suffix)) %>%
    dplyr::select(-.overlap) %>%
    mutate(chrom = str_remove(chrom, "chr"))
  
  
  colnames(df)[colnames(df) %in% c("chrom", "start", "end")] <- old_col_names
  
  return(df)

}

subset_by_significant_gene <- function(df) {
  df %>%
    group_by(FEATURE, GWAS_TYPE) %>%
    filter(P < 0.05 / n())
}
```

## Binning genes (MAGMA) or enhancers (chromMAGMA) to linkage disequilibrium bins
```{r,eval = TRUE}

ld_annotation_originalfeature<- run_intersection(
  gene_file, 
  c("originalfeature_CHR", "originalfeature_START", "originalfeature_STOP"),
  ".LDBin.originalfeature"
)

ld_annotation_genebody <- run_intersection(
  gene_file, 
  c("gene_body_chromosome_name", "gene_body_start_position", "gene_body_end_position"),
  ".LDBin.gene_body"
)

ld_annotated_gene_file <- full_join(
  ld_annotation_genebody, 
  ld_annotation_originalfeature,
  c("ensembl_gene_id", "external_gene_name", "FEATURE", "GWAS_TYPE", "originalfeature", "originalfeature_CHR", "originalfeature_START", "originalfeature_STOP", "NSNPS", "NPARAM", "N", "ZSTAT", "P", "gene_body_chromosome_name", "gene_body_start_position", "gene_body_end_position", "min_Pvalue", "NEG_LOG10P")
)

rm(ld_annotation_genebody, ld_annotation_originalfeature)

head(ld_annotated_gene_file)

#Writing table of Gene_level results that are now binned into LD bins
#saveRDS(ld_annotated_gene_file, "Data/Gene_Level_Bin.4.9.21.rds")
```



```{r, include = FALSE}
#ld_annotated_gene_file %>%
#  mutate(bin_difference = ID.LDBin.originalfeature - ID.LDBin.gene_body) %>%
#  count(bin_difference) %>%
#  filter(bin_difference != 0) %>%
#  ggplot(aes(x = bin_difference, y = n)) +
#  geom_point()
#
#ld_annotated_gene_file %>%
#  subset_by_significant_gene() %>%
#  count(ID.LDBin.gene_body) %>%
#  ungroup() %>%
#  spread(FEATURE, n) %>%
#  rename(LD_bin_number = ID.LDBin.gene_body) %>%
#  print(n = nrow(.))
#
## Ignore below ----
#
#full_join(
#  ld_annotation_genebody, 
#  ld_annotation_originalfeature,
#  c("ensembl_gene_id", "external_gene_name", "FEATURE", "GWAS_TYPE", "originalfeature", "originalfeature_CHR", #"originalfeature_START", "originalfeature_STOP", "NSNPS", "NPARAM", "N", "ZSTAT", "P", "gene_body_chromosome_name", #"gene_body_start_position", "gene_body_end_position", "min_Pvalue", "NEG_LOG10P")
#  ) %>% 
#  filter_all(any_vars(is.na(.))) %>%
#  t()
#
#filter(ld_annotation_genebody, ensembl_gene_id == "ENSG00000117419", GWAS_TYPE == "CCOC", originalfeature == #"1:44820025-44821670") %>%
#  t()
#
#full_join(
#  run_intersection(
#    "gene_body_chromosome_name", 
#    "gene_body_start_position", 
#    "gene_body_end_position",
#    ".LDBin.gene_body"
#  ),
#  
#)
#
#significant_genes_body <- run_intersection(
#  "gene_body_chromosome_name", 
#  "gene_body_start_position", 
#  "gene_body_end_position",
#  ".LDBin.gene_body"
#  ) %>%
#  subset_by_significant_gene() %>%
#  dplyr::select(ensembl_gene_id:originalfeature)
#
#significant_enhancer <- run_intersection(
#  "originalfeature_CHR", 
#  "originalfeature_START", 
#  "originalfeature_STOP"
#) %>%
#  subset_by_significant_gene() %>%
#  dplyr::select(ensembl_gene_id:originalfeature)
#
#run_intersection(
#  "originalfeature_CHR", 
#  "originalfeature_START", 
#  "originalfeature_STOP"
#) %>%
#  subset_by_significant_gene() %>%
#  filter(ensembl_gene_id == "ENSG00000125637") %>%
#  View()
#  group_by(ensembl_gene_id, external_gene_name, FEATURE, GWAS_TYPE, originalfeature) %>%
#  count() %>%
#  filter(n > 1)
#
#full_join(significant_enhancer, significant_genes_body) %>%
#  distinct()
#  filter_all(any_vars(is.na(.)))
```
