---
title: "06 Pathway Gene-Set Enrichment Analysis chromMAGMA"
output: html_document
---
*Robbin Nameki, Anamay Shetty et al.*
*August 2021*

# Introduction
The following script describes the process of conducting gene-set pathway enrichment analysis using the R package ClusterProfiler. This analysis takes a gene-set (in this case pathway gene-sets from the GO data base) and asks whether the gene-set is enriched in a ranked list (in this case chromMAGMA genes ranked by RNA expression weighted P-values). The gseGO() functions takes a bit to run. 

This document can also be adapted to analyze conventional MAGMA gene-lists

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(clusterProfiler)
library(enrichplot)
library(ggplot2)
library(DESeq2)
library(reshape2)
library(biomaRt)
library(UpSetR)
library(ggpubr)
```

## Sourcecode
```{r,eval = TRUE}
source('Rscripts/Utils.R'  )
```

## Getting entrezGeneIDs for the gene-list
The identifiers in the weighted gene-lists are converted to entrezGeneIDs to match the identifiers in the GO database. 
HLA genes are removed to reduce bias to these highly active regions. 
```{r, eval = TRUE}
#Getting EntrezGeneIDs
ensembl <- useEnsembl_GRCh37()
bm <- getBM(attributes=c('ensembl_gene_id', 'external_gene_name', 'entrezgene_id'),
            filters = 'chromosome_name',
            values = c(1:22,'X'),
            mart = ensembl)
bm.unique_external_gene_name <- bm[!duplicated(bm[,'external_gene_name']),]

#Assigning EntrezGeneIDs to analysis
Gene_Level_Weighted_chromMAGMA <- read_gene_level_weighted()
gene_level.ensembl <- merge(Gene_Level_Weighted_chromMAGMA, bm.unique_external_gene_name,by = 'external_gene_name')


gene_level.ensembl.entrez <- gene_level.ensembl[!gene_level.ensembl$entrezgene_id =='',]

#Taking out HLA genes from dataframe
HLA_genes <- read_HLA_genes()
gene_level.ensembl.entrez.noHLA <- gene_level.ensembl.entrez[!(gene_level.ensembl.entrez$external_gene_name %in% HLA_genes$hgnc_symbol),]
```

## Getting Gene Lists
```{r, eval =TRUE}
histotype = c('HGSOC','LGSOC','CCOC','NMOC','MOC','EnOC')
for(i in histotype) {
#chromMAGMA
  assign(paste0('chromMAGMA_',i),
        gene_level.ensembl.entrez.noHLA %>%
           filter(FEATURE == 'chromMAGMA') %>%
           filter(GWAS_TYPE == paste(i))%>%
    dplyr::select(entrezgene_id,weighted_P,external_gene_name) %>%
      arrange(desc(weighted_P))
        )
}



#chromMAGMA_List
chromMAGMA.geneList.CCOC <- chromMAGMA_CCOC[,2]
names(chromMAGMA.geneList.CCOC) = as.character(chromMAGMA_CCOC[,1])
chromMAGMA.geneList.CCOC = sort(chromMAGMA.geneList.CCOC, decreasing = TRUE)

chromMAGMA.geneList.EnOC <- chromMAGMA_EnOC[,2]
names(chromMAGMA.geneList.EnOC) = as.character(chromMAGMA_EnOC[,1])
chromMAGMA.geneList.EnOC = sort(chromMAGMA.geneList.EnOC, decreasing = TRUE)

chromMAGMA.geneList.HGSOC <- chromMAGMA_HGSOC[,2]
names(chromMAGMA.geneList.HGSOC) = as.character(chromMAGMA_HGSOC[,1])
chromMAGMA.geneList.HGSOC = sort(chromMAGMA.geneList.HGSOC, decreasing = TRUE)

chromMAGMA.geneList.LGSOC <- chromMAGMA_LGSOC[,2]
names(chromMAGMA.geneList.LGSOC) = as.character(chromMAGMA_LGSOC[,1])
chromMAGMA.geneList.LGSOC = sort(chromMAGMA.geneList.LGSOC, decreasing = TRUE)

chromMAGMA.geneList.MOC <- chromMAGMA_MOC[,2]
names(chromMAGMA.geneList.MOC) = as.character(chromMAGMA_MOC[,1])
chromMAGMA.geneList.MOC = sort(chromMAGMA.geneList.MOC, decreasing = TRUE)

chromMAGMA.geneList.NMOC <- chromMAGMA_NMOC[,2]
names(chromMAGMA.geneList.NMOC) = as.character(chromMAGMA_NMOC[,1])
chromMAGMA.geneList.NMOC = sort(chromMAGMA.geneList.NMOC, decreasing = TRUE)

```
## GSEGo 
```{r, eval = TRUE, warning = FALSE}
organism = 'org.Hs.eg.db'
#BiocManager::install(organism, character.only = TRUE)
library(organism, character.only = TRUE)
organism.2 = org.Hs.eg.db

chromMAGMA.CCOC.gse <- gseGO(geneList= chromMAGMA.geneList.CCOC, 
             ont ="ALL", 
             keyType = "ENTREZID", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism.2, 
             pAdjustMethod = "BH")

chromMAGMA.EnOC.gse <- gseGO(geneList= chromMAGMA.geneList.EnOC, 
             ont ="ALL", 
             keyType = "ENTREZID", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism.2, 
             pAdjustMethod = "BH")

chromMAGMA.HGSOC.gse <- gseGO(geneList= chromMAGMA.geneList.HGSOC, 
             ont ="ALL", 
             keyType = "ENTREZID", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism.2, 
             pAdjustMethod = "BH")

chromMAGMA.LGSOC.gse <- gseGO(geneList= chromMAGMA.geneList.LGSOC, 
             ont ="ALL", 
             keyType = "ENTREZID", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism.2, 
             pAdjustMethod = "BH")

chromMAGMA.MOC.gse <- gseGO(geneList= chromMAGMA.geneList.MOC, 
             ont ="ALL", 
             keyType = "ENTREZID", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism.2, 
             pAdjustMethod = "BH")

chromMAGMA.NMOC.gse <- gseGO(geneList= chromMAGMA.geneList.NMOC, 
             ont ="ALL", 
             keyType = "ENTREZID", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism.2, 
             pAdjustMethod = "BH")

##output saved to one file.
##rbinding these dataframes makes ST4
#saveRDS(chromMAGMA.CCOC.gse, 'Data/GSEgo')
#saveRDS(chromMAGMA.EnOC.gse, 'Data/GSEgo')
#saveRDS(chromMAGMA.HGSOC.gse, 'Data/GSEgo')
#saveRDS(chromMAGMA.LGSOC.gse, 'Data/GSEgo')
#saveRDS(chromMAGMA.MOC.gse, 'Data/GSEgo')
#saveRDS(chromMAGMA.NMOC.gse, 'Data/GSEgo')
```


```{r,eval=TRUE}
file_location <- "Data/GSEgo"

format_output_df <- function(file_location) {
  as_tibble(read_rds(file_location)@result) %>%
    separate_rows(core_enrichment, sep = "/") %>%
    dplyr::select("ENTREZEGENE_ID" = core_enrichment, Description, ID, setSize, NES, p.adjust, ONTOLOGY)
}

tibble(histotype = str_remove(list.files(file_location), ".gse.RDS"), 
  file_location = list.files(file_location, full.names = TRUE)) %>%
  group_by(histotype) %>%
  dplyr::mutate(results = map(file_location, format_output_df)) %>%
  dplyr::select(-file_location) %>%
  unnest(results) %>%
  write_tsv("Data/cleaned_pathway_gsea_output.5.25.21.txt")
```

#Histotype and pathway-specific analysis
```{r, eval = TRUE}
#getting list for upset
chromMAGMA.CCOC.gse.df <- as.data.frame(chromMAGMA.CCOC.gse@result) %>%
  filter(setSize >= 25)
chromMAGMA.CCOC.gse.df$GWAS_TYPE <- 'CCOC'
chromMAGMA.EnOC.gse.df <- as.data.frame(chromMAGMA.EnOC.gse@result)%>%
  filter(setSize >= 25)
chromMAGMA.EnOC.gse.df$GWAS_TYPE <- 'EnOC'
chromMAGMA.HGSOC.gse.df <- as.data.frame(chromMAGMA.HGSOC.gse@result)%>%
  filter(setSize >= 25)
chromMAGMA.HGSOC.gse.df$GWAS_TYPE <- 'HGSOC'
chromMAGMA.LGSOC.gse.df <- as.data.frame(chromMAGMA.LGSOC.gse@result)%>%
  filter(setSize >= 25)
chromMAGMA.LGSOC.gse.df$GWAS_TYPE <- 'LGSOC'
chromMAGMA.MOC.gse.df <- as.data.frame(chromMAGMA.MOC.gse@result)%>%
  filter(setSize >= 25)
chromMAGMA.MOC.gse.df$GWAS_TYPE <- 'MOC'
chromMAGMA.NMOC.gse.df <- as.data.frame(chromMAGMA.NMOC.gse@result)%>%
  filter(setSize >= 25)
chromMAGMA.NMOC.gse.df$GWAS_TYPE <- 'NMOC'

#ST4
ST4 <- rbind(chromMAGMA.CCOC.gse.df,
      chromMAGMA.EnOC.gse.df,
      chromMAGMA.HGSOC.gse.df,
      chromMAGMA.LGSOC.gse.df,
      chromMAGMA.MOC.gse.df,
      chromMAGMA.NMOC.gse.df)
head(ST4)

pathway.list <- list(
  CCOC = chromMAGMA.CCOC.gse.df$Description,
  EnOC = chromMAGMA.EnOC.gse.df$Description,
  HGSOC = chromMAGMA.HGSOC.gse.df$Description,
  LGSOC = chromMAGMA.LGSOC.gse.df$Description,
  MOC = chromMAGMA.MOC.gse.df$Description
)

#upset plot
library(UpSetR)
upset(fromList(pathway.list),
      nsets = length(pathway.list),
      #$empty.intersections = 'on',
      order.by ='freq')

#extracting binary results within upset plot
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
annotated_pathway <- fromList(pathway.list)
annotated_pathway$Description <- rownames(annotated_pathway)
annotated_pathway$binary <-  paste(annotated_pathway$CCOC,
                            annotated_pathway$EnOC,
                            annotated_pathway$HGSOC,
                            annotated_pathway$LGSOC,
                            annotated_pathway$MOC) 

annotated_pathway$anno <- ifelse(annotated_pathway$binary %in% "1 0 0 0 0",'CCOC Specific',
                             ifelse(annotated_pathway$binary %in% "0 1 0 0 0",'EnOC Specific',
                                    ifelse(annotated_pathway$binary %in% "0 0 1 0 0",'HGSOC Specific',
                                                  ifelse(annotated_pathway$binary %in% "0 0 0 1 0",'LGSOC Specific',
                                                         ifelse(annotated_pathway$binary %in% "0 0 0 0 1",'MOC Specific',
                                                                ifelse(annotated_pathway$binary %in% "1 1 1 1 1",'All Common','NA'))))))

annotated_pathway <- annotated_pathway %>% filter(!anno == 'NA')
annotated_pathway<- annotated_pathway[order(annotated_pathway$anno),]



#dotplot
transcription <- annotated_pathway %>%
  filter(anno == 'All Common') %>%
    filter(str_detect(Description,"transcription"))

splicing <- annotated_pathway %>%
  filter(anno == 'All Common') %>%
     filter(str_detect(Description,"splic"))

chromatin <- annotated_pathway %>%
  filter(anno == 'All Common') %>%
     filter(str_detect(Description,"chromatin"))
dotpot_Des <- rbind(transcription,splicing,chromatin)

all_pathway <- rbind(as.data.frame(chromMAGMA.CCOC.gse@result) %>% mutate(histotype = 'CCOC'),
                     as.data.frame(chromMAGMA.EnOC.gse@result) %>% mutate(histotype = 'EnOC'),
                     as.data.frame(chromMAGMA.HGSOC.gse@result) %>% mutate(histotype = 'HGSOC'),
                     as.data.frame(chromMAGMA.LGSOC.gse@result) %>% mutate(histotype = 'LGSOC'),
                     as.data.frame(chromMAGMA.MOC.gse@result) %>% mutate(histotype = 'MOC')
      ) %>%
dplyr::select(histotype,Description, p.adjust,NES) %>%
mutate(NEG_LOG10P = -log10(p.adjust)) %>%
filter(Description %in% dotpot_Des$Description) %>%
group_by(histotype) %>%
dplyr::select(histotype,Description,NEG_LOG10P, NES)

ggplot(all_pathway,aes(factor(histotype),Description)) + 
  geom_point(aes(colour=NEG_LOG10P,size=NES)) + 
  scale_colour_gradient(low="blue", high="red") + 
  theme_bw() + 
  theme(axis.text=element_text(size=7), 
        axis.title=element_text(size=7,face="bold")) + 
  theme(axis.text.x = element_text(size = 12, angle = 90)) + 
  scale_size_continuous(range = c(0,6))
```

#Figure 3b
```{r,eval = TRUE, fig.width=10,fig.height=11}
CCOC_enriched <- annotated_pathway %>%
  filter(anno == 'CCOC Specific')
chromMAGMA.CCOC.enriched <- chromMAGMA.CCOC.gse.df %>%
  filter(Description %in% CCOC_enriched$Description) %>%
  mutate(NEG_LOG10_P = -log10(p.adjust)) %>% 
  top_n(10, NES)
chromMAGMA.CCOC.enriched$Description <- str_trunc(chromMAGMA.CCOC.enriched$Description, 50) 
chromMAGMA.CCOC.enriched$Description <- factor(chromMAGMA.CCOC.enriched$Description, levels=chromMAGMA.CCOC.enriched[order(chromMAGMA.CCOC.enriched$NEG_LOG10_P,decreasing=T),]$Description)

CCOC.plot <- ggplot(chromMAGMA.CCOC.enriched,aes(x = NEG_LOG10_P,
                                                 y = Description)) + 
  geom_col(fill = 'red') +
  coord_flip()+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  xlim(0,2.5)

#EnOC
EnOC_enriched <- annotated_pathway %>%
  filter(anno == 'EnOC Specific')
chromMAGMA.EnOC.enriched <- chromMAGMA.EnOC.gse.df %>%
  filter(Description %in% EnOC_enriched$Description) %>%
  mutate(NEG_LOG10_P = -log10(p.adjust)) %>% 
  top_n(10, NES) 
chromMAGMA.EnOC.enriched$Description <- str_trunc(chromMAGMA.EnOC.enriched $Description, 50) 
chromMAGMA.EnOC.enriched$Description <- factor(chromMAGMA.EnOC.enriched $Description, levels=chromMAGMA.EnOC.enriched[order(chromMAGMA.EnOC.enriched $NEG_LOG10_P,decreasing=T),]$Description)

EnOC.plot <- ggplot(chromMAGMA.EnOC.enriched ,
                    aes(x = NEG_LOG10_P,
                        y = Description)) + 
  geom_col(fill = 'blue')+
  coord_flip()+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())+
  xlim(0,2.5)

#HGSOC
HGSOC_enriched <- annotated_pathway %>%
  filter(anno == 'HGSOC Specific')
chromMAGMA.HGSOC.enriched <- chromMAGMA.HGSOC.gse.df %>%
  filter(Description %in% HGSOC_enriched$Description) %>%
  mutate(NEG_LOG10_P = -log10(p.adjust)) %>% 
  top_n(10, NES)
chromMAGMA.HGSOC.enriched$Description <- str_trunc(chromMAGMA.HGSOC.enriched$Description, 50) 
chromMAGMA.HGSOC.enriched$Description <- factor(chromMAGMA.HGSOC.enriched$Description, levels=chromMAGMA.HGSOC.enriched[order(chromMAGMA.HGSOC.enriched$NEG_LOG10_P,decreasing=T),]$Description)

HGSOC.plot <- ggplot(chromMAGMA.HGSOC.enriched,
                      aes(x = NEG_LOG10_P, 
                          y = Description)) +
  geom_col(fill = 'green') +
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())+
  xlim(0,2.5)
#LGSOC
LGSOC_enriched <- annotated_pathway %>%
  filter(anno == 'LGSOC Specific')
chromMAGMA.LGSOC.enriched <- chromMAGMA.LGSOC.gse.df %>%
  filter(Description %in% LGSOC_enriched$Description) %>%
  mutate(NEG_LOG10_P = -log10(p.adjust)) %>% 
  top_n(10, NES)
chromMAGMA.LGSOC.enriched$Description <- str_trunc(chromMAGMA.LGSOC.enriched$Description, 50) 
chromMAGMA.LGSOC.enriched <- chromMAGMA.LGSOC.enriched %>%
  distinct(Description, .keep_all = TRUE)

chromMAGMA.LGSOC.enriched$Description <- factor(chromMAGMA.LGSOC.enriched$Description, levels=chromMAGMA.LGSOC.enriched[order(chromMAGMA.LGSOC.enriched$NEG_LOG10_P,decreasing=T),]$Description)

LGSOC.plot <- ggplot(chromMAGMA.LGSOC.enriched,
                     aes(x = NEG_LOG10_P,
                         y = Description)) + 
  geom_col(fill = 'orange')+
  coord_flip()+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())+
  xlim(0,2.5)

#MOC
MOC_enriched <- annotated_pathway %>%
  filter(anno == 'MOC Specific')
chromMAGMA.MOC.enriched <- chromMAGMA.MOC.gse.df %>%
  filter(Description %in% MOC_enriched$Description) %>%
  mutate(NEG_LOG10_P = -log10(p.adjust)) %>% 
  top_n(10, NES)
chromMAGMA.MOC.enriched$Description <- str_trunc(chromMAGMA.MOC.enriched$Description, 50) 
chromMAGMA.MOC.enriched$Description <- factor(chromMAGMA.MOC.enriched$Description, levels=chromMAGMA.MOC.enriched[order(chromMAGMA.MOC.enriched$NEG_LOG10_P,decreasing=T),]$Description)

MOC.plot <- ggplot(chromMAGMA.MOC.enriched,
                   aes(x = NEG_LOG10_P,
                       y = Description)) + 
  geom_bar(stat="identity", fill = 'purple')+
  coord_flip()+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())+
  xlim(0,2.5)

ggarrange(CCOC.plot,
          EnOC.plot,
          HGSOC.plot,
          LGSOC.plot,
          MOC.plot,
          ncol = 5,
          nrow = 1)
```


```{r,include = FALSE}
## Descriptive Statistics of chromMAGMA Pathway Analysis
df <- read.table('Data/cleaned_pathway_gsea_output.5.25.21.txt', sep = '\t', header = TRUE, quote = "")
df <- df %>%
  dplyr::rename(entrezgene_id = ENTREZEGENE_ID) %>%
  filter(NES > 0)
df.symbol <- merge(df,bm.unique_external_gene_name,by = 'entrezgene_id', all.x = TRUE)

df.symbol.CCOC <- df.symbol %>%
  filter(histotype == 'CCOC') %>%
  distinct(Description, .keep_all = TRUE)

df.symbol.EnOC <- df.symbol %>%
  filter(histotype == 'EnOC') %>%
  distinct(Description, .keep_all = TRUE)

df.symbol.HGSOC <- df.symbol %>%
  filter(histotype == 'HGSOC') %>%
  distinct(Description, .keep_all = TRUE)

df.symbol.LGSOC <- df.symbol %>%
  filter(histotype == 'LGSOC') %>%
  distinct(Description, .keep_all = TRUE)

df.symbol.MOC <- df.symbol %>%
  filter(histotype == 'MOC') %>%
  distinct(Description, .keep_all = TRUE)

df.symbol.NMOC <- df.symbol %>%
  filter(histotype == 'NMOC') %>%
  distinct(Description, .keep_all = TRUE)

df.symbol.list <- c('df.symbol.CCOC','df.symbol.EnOC','df.symbol.HGSOC','df.symbol.LGSOC','df.symbol.MOC','df.symbol.NMOC')

#Upset
df.upset <- list(
  CCOC = df.symbol.CCOC$Description,
  EnOC = df.symbol.EnOC$Description,
  HGSOC = df.symbol.HGSOC$Description,
  LGSOC = df.symbol.LGSOC$Description,
  MOC = df.symbol.MOC$Description,
  NMOC = df.symbol.NMOC$Description
)

plot <- upset(fromList(df.upset), nsets = length(df.upset), 
              empty.intersections = "on", 
              order.by = "freq", 
              point.size = 3.5, 
              text.scale = 1.5,
              line.size = 2)
#finding common pathways
df.inner <- inner_join(df.symbol.CCOC, df.symbol.EnOC, by='Description') %>%
  inner_join(.,df.symbol.HGSOC, by='Description') %>%
  inner_join(.,df.symbol.LGSOC, by='Description')%>%
  inner_join(.,df.symbol.MOC, by='Description')%>%
  inner_join(.,df.symbol.NMOC, by='Description')
length(unique(df.inner$Description))
```

## Getting more info on what pathways are common 
```{r, include = FALSE}
## Finding Genes involved in Splicing
df.acrosshistotypes <- df.symbol[df.symbol$Description %in% df.inner$Description,] 
##All Splice
splicing <- df.acrosshistotypes %>%
filter(str_detect(Description,'splic'))
splicing <- splicing %>%
  distinct(Description, .keep_all = TRUE) 

splicing.genes <- df.acrosshistotypes %>%
  filter(str_detect(Description,'splic')) %>%
  distinct(external_gene_name, .keep_all = TRUE) %>%
  dplyr::select(ensembl_gene_id,external_gene_name,ID,Description)
splicing.genes$gene_biotype <- 'splicing.genes'

## all transcription
transcription <- df.acrosshistotypes %>%
filter(str_detect(Description,'transcription factor'))
transcription <- transcription %>%
  distinct(Description, .keep_all = TRUE) 

transcription.genes <- df.acrosshistotypes %>%
  filter(str_detect(Description,'transcription')) %>%
  distinct(external_gene_name, .keep_all = TRUE) %>%
  dplyr::select(ensembl_gene_id,external_gene_name,ID,Description)
transcription.genes$gene_biotype <- 'transcription.genes'

splicing_transcription_pathway_genes_common_all_histotypes <- rbind(splicing.genes,transcription.genes)

## chromatin?
chromatin <- df.acrosshistotypes %>%
filter(str_detect(Description,'chromatin'))
chromatin <- chromatin %>%
  distinct(Description, .keep_all = TRUE) 

chromatin.genes <- df.acrosshistotypes %>%
  filter(str_detect(Description,'chromatin')) %>%
  distinct(external_gene_name, .keep_all = TRUE) %>%
  dplyr::select(ensembl_gene_id,external_gene_name,ID,Description)
chromatin.genes$gene_biotype <- 'chromatin.genes'

## cell cycle?
cycle <- df.acrosshistotypes %>%
filter(str_detect(Description,'cycle'))
cycle <- cycle %>%
  distinct(Description, .keep_all = TRUE) 

cycle.genes <- df.acrosshistotypes %>%
  filter(str_detect(Description,'cycle')) %>%
  distinct(external_gene_name, .keep_all = TRUE) %>%
  dplyr::select(ensembl_gene_id,external_gene_name,ID,Description)
cycle.genes$gene_biotype <- 'cycle.genes'

```


```{r, include = FALSE}
## Figure 2D chromMAGMA GSEA Dotplots
pathway_list <- as.list(
df.inner %>%
  filter(str_detect(Description,c('transcription factor','mRNA'))) %>%
  dplyr::select(Description))


all_pathway <- rbind(as.data.frame(chromMAGMA.CCOC.gse@result) %>% mutate(histotype = 'CCOC'),
                     as.data.frame(chromMAGMA.EnOC.gse@result) %>% mutate(histotype = 'EnOC'),
                     as.data.frame(chromMAGMA.HGSOC.gse@result) %>% mutate(histotype = 'HGSOC'),
                     as.data.frame(chromMAGMA.LGSOC.gse@result) %>% mutate(histotype = 'LGSOC'),
                     as.data.frame(chromMAGMA.MOC.gse@result) %>% mutate(histotype = 'MOC'),
                     as.data.frame(chromMAGMA.NMOC.gse@result) %>% mutate(histotype = 'NMOC')
      ) %>%
dplyr::select(histotype,Description, p.adjust,NES) %>%
mutate(NEG_LOG10P = -log10(p.adjust)) %>%
group_by(histotype) %>%
filter(str_detect(Description,"transcription factor|splicing")) %>%
dplyr::select(histotype,Description,NEG_LOG10P, NES)



ggplot(all_pathway,aes(factor(histotype),Description)) + 
  geom_point(aes(colour=NEG_LOG10P,size=NES)) + 
  scale_colour_gradient(low="blue", high="red") + 
  scale_size() + 
  theme_bw() + 
  theme(axis.text=element_text(size=7), 
        axis.title=element_text(size=7,face="bold")) + 
  theme(axis.text.x = element_text(size = 12, angle = 90))
```


```{r,include = FALSE}
#Making plots showing where the splicing genes in pathway are ranked in chromMAGMA
gene_level_weighted <- read_gene_level_weighted()

transcription.genes.path <- transcription.genes %>%
  filter(Description == 'transcription factor binding')
list <- data.frame(GENES = c('ABLIM3',
'ACTN1',
'AHCYL1',
'ALCAM',
'ATP1B4',
'ATP5MC1',
'BDNF',
'BEND4',
'BHLHE22',
'BNC2',
'CALD1',
'CAMK1D',
'CCDC24',
'CFAP20',
'CLTC',
'CNTN6',
'DIO2',
'DLG2',
'DNAI1',
'ELK4',
'ELMO3',
'EML1',
'FAM172A',
'FAM219A',
'FAM71F1',
'FEZF2',
'GABRA1',
'GABRG2',
'GGN',
'GPC4',
'GPR27',
'HIVEP1',
'HNRNPA0',
'HOXA9',
'HOXB4',
'HOXB5',
'HOXB7',
'HOXB8',
'HOXD12',
'HOXD8',
'HPN',
'HR',
'IARS1',
'IGSF22',
'JMJD1C',
'KCNQ1DN',
'KLHL13',
'LHFPL1',
'LRMDA',
'LRRC74A',
'LUC7L3',
'MAN1C1',
'MAP2K5',
'MARCHF1',
'METTL26',
'MNT',
'MYBPC1',
'MYF5',
'NABP2',
'NECAB3',
'NEUROD6',
'NEUROG1',
'NOL4',
'NPM3',
'NR1D1',
'NRG1',
'NRGN',
'OLFM2',
'OSBPL7',
'PAK1IP1',
'PAPPA',
'PDZD7',
'PIK3R3',
'POU4F3',
'PPM1A',
'PPP2R2B',
'PTCH1',
'PUM2',
'PURA',
'RARB',
'RBM39',
'RELCH',
'RERG',
'RGS12',
'RHOBTB1',
'RHOQ',
'RNF220',
'SH3GL3',
'SLC25A28',
'SP6',
'SRPK2',
'SUPT16H',
'TAF5',
'TANK',
'TCF7L2',
'TEK',
'TIA1',
'TIAF1',
'TMEM255A',
'TMTC2',
'TRERF1',
'TSHZ2',
'TSPAN2',
'TWIST1',
'WNT4',
'ZNF385B'))

list <- data.frame(GENES = c('HOXB5',
'HOXB7',
'HOXB8',
'LUC7L3',
'MAP2K5',
'NEUROD6',
'RBM39',
'SRPK2'))
gene_level_weighted %>%
  group_by(GWAS_TYPE) %>%
  mutate(rank_weighted_P = min_rank(-weighted_P)) %>%
  filter(GWAS_TYPE == 'NMOC') %>%
  filter(FEATURE == c('chromMAGMA')) %>%
  {ggplot(., aes(x = rank_weighted_P, y = weighted_P)) + 
      geom_line() + 
      #geom_hline(yintercept = c(-1.96, 1.96), alpha = 0.3, linetype = "dashed") +
      #geom_point(data = filter(., hgnc_symbol %in% list$GENES), colour = "black", size = 2) +
      geom_point(data = filter(., external_gene_name %in% list$GENES), colour = "#4AAE42", size = 3) +  
      ggrepel::geom_text_repel(data = filter(., external_gene_name %in% list$GENES), 
                               aes(label = external_gene_name), 
                               max.overlaps = 20,
                               size = 3,
                               seed= 100) +
      facet_grid(rows = vars(FEATURE), cols = vars(GWAS_TYPE)) +
      xlab("-log10(weighted_Pvalue) Rank") + ylab("-log10(Weighted_Pvalue)") +
      #ggtitle("Ranked Z-score for Selected Genes", subtitle = "Gene-Centric Analysis") +
      scale_color_brewer(palette = "Dark2") +
      theme_minimal()
  }

```
