---
title: "08 Super-Enhancer Associated Transcription Graphs"
output: html_document
---
*Robbin Nameki, Anamay Shetty et al.*
*August 2021*

# Introduction
This document is for the visualization (Figure 3A and 3B) of the Super-enhancer associated transcription factor GSEA analysis. Although BROAD GSEA was used for the analysis, the package fgsea was used for the enrichment plot visualization. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sourcecode
```{r,eval = TRUE}
source('Rscripts/Utils.R')
```

## Libraries
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(fgsea)
library(ggpubr)
```

## Gene-Set 
The gene-set in this analysis are super-enhancer associated genes from Corona et al. 2020
```{r, eval = TRUE}
geneset <- read.table('Data/SuperEnhancer_TFs.txt', sep = '\t', header = TRUE)

geneset_nested <- geneset %>%
  group_by(histotype)  %>%
  split(f = as.factor(.$histotype))

```

## Making GSEA Plots chromMAGMA
```{r,eval = TRUE}
#Making nested dataframe chromMAGMA
gene_file <- read_gene_level_weighted()

gene_file_nested <- gene_file %>%
  filter(FEATURE == 'chromMAGMA') %>%
  dplyr::select(GWAS_TYPE,external_gene_name,weighted_P) %>%
  group_by(GWAS_TYPE)  %>%
  split(f = as.factor(.$GWAS_TYPE))

#naming weighted_P chromMAGMA
names(gene_file_nested$CCOC$weighted_P) <- gene_file_nested$CCOC$external_gene_name
names(gene_file_nested$EnOC$weighted_P) <- gene_file_nested$EnOC$external_gene_name
names(gene_file_nested$HGSOC$weighted_P) <- gene_file_nested$HGSOC$external_gene_name
names(gene_file_nested$LGSOC$weighted_P) <- gene_file_nested$LGSOC$external_gene_name
names(gene_file_nested$MOC$weighted_P) <- gene_file_nested$MOC$external_gene_name
names(gene_file_nested$NMOC$weighted_P) <- gene_file_nested$NMOC$external_gene_name

chromMAGMA.CCOC <- plotEnrichment(geneset_nested$CCOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested$CCOC$weighted_P) + 
  labs(title="") +
  labs(title="CCOC") +
  geom_line(color="red")

chromMAGMA.EnOC <- plotEnrichment(geneset_nested$EnOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested$EnOC$weighted_P) + 
  labs(title="") +
  labs(title="EnOC") +
  geom_line(color="blue")

chromMAGMA.HGSOC <- plotEnrichment(geneset_nested$HGSOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested$HGSOC$weighted_P) + 
  labs(title="") +
  labs(title="HGSOC") +
  geom_line(color="green")

chromMAGMA.MOC <- plotEnrichment(geneset_nested$MOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested$MOC$weighted_P) + 
  labs(title="") +
  labs(title="MOC") +
  geom_line(color="purple")

```

## Making GSEA Plots MAGMA
```{r,eval = TRUE}
gene_file_MAGMA <- read_gene_level()
#Making nested dataframe MAGMA
gene_file_nested_MAGMA <- gene_file_MAGMA %>%
  filter(FEATURE == 'MAGMA') %>%
  dplyr::select(GWAS_TYPE,external_gene_name,NEG_LOG10P) %>%
  group_by(GWAS_TYPE)  %>%
  split(f = as.factor(.$GWAS_TYPE))

#naming NEG_LOG10P MAGMA
names(gene_file_nested_MAGMA$CCOC$NEG_LOG10P) <- gene_file_nested_MAGMA$CCOC$external_gene_name
names(gene_file_nested_MAGMA$EnOC$NEG_LOG10P) <- gene_file_nested_MAGMA$EnOC$external_gene_name
names(gene_file_nested_MAGMA$HGSOC$NEG_LOG10P) <- gene_file_nested_MAGMA$HGSOC$external_gene_name
names(gene_file_nested_MAGMA$LGSOC$NEG_LOG10P) <- gene_file_nested_MAGMA$LGSOC$external_gene_name
names(gene_file_nested_MAGMA$MOC$NEG_LOG10P) <- gene_file_nested_MAGMA$MOC$external_gene_name
names(gene_file_nested_MAGMA$NMOC$NEG_LOG10P) <- gene_file_nested_MAGMA$NMOC$external_gene_name

MAGMA.CCOC <- plotEnrichment(geneset_nested$CCOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested_MAGMA$CCOC$NEG_LOG10P) + 
  labs(title="") +
  labs(title="CCOC") +
  geom_line(color="red")

MAGMA.EnOC <- plotEnrichment(geneset_nested$EnOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested_MAGMA$EnOC$NEG_LOG10P) + 
  labs(title="") +
  labs(title="EnOC") +
  geom_line(color="blue")

MAGMA.HGSOC <- plotEnrichment(geneset_nested$HGSOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested_MAGMA$HGSOC$NEG_LOG10P) + 
  labs(title="") +
  labs(title="HGSOC") +
  geom_line(color="green")

MAGMA.MOC <- plotEnrichment(geneset_nested$MOC$OVERLAP_PROXIMAL_CLOSEST_GENES,gene_file_nested_MAGMA$MOC$NEG_LOG10P) + 
  labs(title="") +
  labs(title="MOC") +
  geom_line(color="purple")
```
## Putting together all the enrichment plots 
```{r,eval = TRUE}
ggarrange(
  chromMAGMA.CCOC, 
  chromMAGMA.EnOC, 
  chromMAGMA.HGSOC,
  chromMAGMA.MOC,
    MAGMA.CCOC, 
    MAGMA.EnOC,
    MAGMA.HGSOC,
    MAGMA.MOC,
  ncol = 4,
  nrow = 2
)
```
## Leading Edge Genes and Plotting
The following segments extracts the genes on the leading edge of the BROAD GSEA analysis from 07.5 and plots their ranking on the chromMAGMA gene-list

### Extracting Leading Edge Genes
This generates supplementary table 5
```{r,eval = TRUE}
myTF.CCOC.top10 <- read.delim('Data/SE_TF_LeadingEdge_Genes/CCOC_SUPERENHANCERS.xls')
myTF.CCOC.top10 <- myTF.CCOC.top10 %>%
  filter(CORE.ENRICHMENT == 'Yes') %>%
  dplyr::rename(gene_id = PROBE)
myTF.CCOC.top10$GWAS_TYPE <- 'clearcell'

myTF.EnOC.top10 <- read.delim('Data/SE_TF_LeadingEdge_Genes/ENOC_SUPERENHANCERS.xls')
myTF.EnOC.top10 <- myTF.EnOC.top10 %>%
  filter(CORE.ENRICHMENT == 'Yes') %>%
  dplyr::rename(gene_id = PROBE)
myTF.EnOC.top10$GWAS_TYPE <- 'endometrioid'

myTF.HGSOC.top10 <- read.delim('Data/SE_TF_LeadingEdge_Genes/HGSOC_SUPERENHANCERS.xls')
myTF.HGSOC.top10 <- myTF.HGSOC.top10 %>%
  filter(CORE.ENRICHMENT == 'Yes') %>%
  dplyr::rename(gene_id = PROBE)
myTF.HGSOC.top10$GWAS_TYPE <- 'serous_hg_extra'

myTF.MOC.top10 <- read.delim('Data/SE_TF_LeadingEdge_Genes/MOC_SUPERENHANCERS.xls')
myTF.MOC.top10 <- myTF.MOC.top10 %>%
  filter(CORE.ENRICHMENT == 'Yes') %>%
  dplyr::rename(gene_id = PROBE)
myTF.MOC.top10$GWAS_TYPE <- 'mucinous_all'

LE_list <- rbind(myTF.CCOC.top10[,c(2,5,6,7,10)],
                 myTF.EnOC.top10[,c(2,5,6,7,10)],
                 myTF.HGSOC.top10[,c(2,5,6,7,10)],
                 myTF.MOC.top10[,c(2,5,6,7,10)])
head(LE_list)

#will be used in later analysis
#saveRDS(x = LE_list, file = 'Data/se_tf_le_list.rds')
```


## Leading-Edge plot
Plot showing the ranking of known super-enhancer associated, lineage-specifying transcription factors on the leading edge of the BROAD GSEA analsis relative to gene-list ranking.
```{r, eval = TRUE}
list <- data.frame(GENES = c('WT1','SOX17','PAX8','ESR1','MECOM','NR2F6'))
ccoclist <- data.frame(GENES = c('HNF1B'))
#enoclist <- data.frame(GENES = c('MYC','HIF1A','ESR1','SOX17','RXRA',''))
#moclist <- data.frame(GENES = c('ONECUT2','MYRF'))


hgsocprimarytissueplot <- gene_file %>%
  group_by(GWAS_TYPE) %>%
  mutate(rank_weighted_P = min_rank(-weighted_P)) %>%
  filter(GWAS_TYPE == 'HGSOC') %>%
  filter(FEATURE == c('chromMAGMA')) %>%
  {ggplot(., aes(x = rank_weighted_P, y = weighted_P)) + 
      geom_line() + 
      #geom_hline(yintercept = c(-1.96, 1.96), alpha = 0.3, linetype = "dashed") +
      #geom_point(data = filter(., hgnc_symbol %in% list$GENES), colour = "black", size = 2) +
      geom_point(data = filter(., external_gene_name %in% list$GENES), colour = "#4AAE42", size = 3) +  
      ggrepel::geom_text_repel(data = filter(., external_gene_name %in% list$GENES), 
                               aes(label = external_gene_name), 
                               max.overlaps = 20,
                               size = 3,
                               seed= 100) +
      facet_grid(rows = vars(FEATURE), cols = vars(GWAS_TYPE)) +
      xlab("-log10(weighted_Pvalue) Rank") + ylab("-log10(Weighted_Pvalue)") +
      #ggtitle("Ranked Z-score for Selected Genes", subtitle = "Gene-Centric Analysis") +
      scale_color_brewer(palette = "Dark2") +
      theme_minimal()
  }

ccocprimarytissueplot <- gene_file %>%
  group_by(GWAS_TYPE) %>%
  mutate(rank_weighted_P = min_rank(-weighted_P)) %>%
  filter(GWAS_TYPE == 'CCOC') %>%
  filter(FEATURE == c('chromMAGMA')) %>%
  {ggplot(., aes(x = rank_weighted_P, y = weighted_P)) + 
      geom_line() + 
      #geom_hline(yintercept = c(-1.96, 1.96), alpha = 0.3, linetype = "dashed") +
      #geom_point(data = filter(., hgnc_symbol %in% list$GENES), colour = "black", size = 2) +
      geom_point(data = filter(., external_gene_name %in% ccoclist$GENES), colour = "#E61B1C", size = 3) +  
      ggrepel::geom_text_repel(data = filter(., external_gene_name %in% ccoclist$GENES), 
                               aes(label = external_gene_name), 
                               max.overlaps = 20,
                               size = 3,
                               seed= 100) +
      facet_grid(rows = vars(FEATURE), cols = vars(GWAS_TYPE)) +
      xlab("-log10(weighted_Pvalue) Rank") + ylab("-log10(Weighted_Pvalue)") +
      #ggtitle("Ranked Z-score for Selected Genes", subtitle = "Gene-Centric Analysis") +
      scale_color_brewer(palette = "Dark2") +
      theme_minimal()
  }


ggarrange(
  hgsocprimarytissueplot,
  ccocprimarytissueplot
)
```


