---
title: "10 Getting List of Nexus Transccription Factors"
output: html_document
---
*Robbin Nameki, Anamay Shetty et al.*
*August 2021*
# Introduction
This section introduces the process of identifying Nexus Transcription Factors. Used for the generation of Supplementary Table 7. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Libraries
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

## Sourcecode
```{r,eval = TRUE}
source('Rscripts/Utils.R')
```
## Data
This analysis compares  MsigDB BROAD GSEA results x the transcription factors on the leading edge of the super-enhancer associated GSEA analysis.
```{r, eval = TRUE}
legacy_tf <- getting_legacy_tf_MsigDBGSEA()

histotype = c('HGSOC','LGSOC','CCOC','NMOC','MOC','EnOC')
for(i in histotype) {
  #MAGMA for each histotype
  assign(paste0('MAGMA_sig_',i),
         legacy_tf %>%
           filter(FEATURE == 'MAGMA') %>%
           filter(GWAS_TYPE == paste(i)) %>%
           filter(!is.na(GENE)) %>%
           group_by(FEATURE, GWAS_TYPE) %>%
           filter(NOM.p.val < 0.05) %>%
           filter(FDR.q.val < 0.25)
         ) 
  
  #chromMAGMA for each histotype
  assign(paste0('chromMAGMA_sig_',i),
         legacy_tf %>%
           filter(FEATURE == 'chromMAGMA') %>%
           filter(GWAS_TYPE == paste(i)) %>%
           filter(!is.na(GENE)) %>%
           group_by(FEATURE, GWAS_TYPE) %>%
           filter(NOM.p.val < 0.05) %>%
           filter(FDR.q.val < 0.25)
         )
}

LE_list <- getting_LE_list()
LE_list_nested <- LE_list %>%
  dplyr::rename(GENE = gene_id) %>%
  split(f = as.factor(.$GWAS_TYPE)) 
```

##List
of Nexus TFs (Supplementary Table 7)
```{r,eval= TRUE}
#CCOC
chromMAGMA_tf_CCOC <- chromMAGMA_sig_CCOC[chromMAGMA_sig_CCOC$GENE %in% LE_list_nested$clearcell$GENE,]
chromMAGMA_tf_EnOC <- chromMAGMA_sig_EnOC[chromMAGMA_sig_EnOC$GENE %in% LE_list_nested$endometrioid$GENE,]
chromMAGMA_tf_HGSOC <- chromMAGMA_sig_HGSOC[chromMAGMA_sig_HGSOC$GENE %in% LE_list_nested$serous_hg_extra$GENE,]
chromMAGMA_tf_MOC <- chromMAGMA_sig_MOC[chromMAGMA_sig_MOC$GENE %in% LE_list_nested$mucinous_all$GENE,]


LE_list2 <- LE_list %>%
  filter(!GWAS_TYPE == 'mucinous_all') %>%
  dplyr::rename(GENE = gene_id)
chromMAGMA_tf_NMOC <- chromMAGMA_sig_NMOC[chromMAGMA_sig_NMOC$GENE %in% LE_list2$GENE,]

tf_list_final <- rbind(chromMAGMA_tf_CCOC,
      chromMAGMA_tf_EnOC,
      chromMAGMA_tf_HGSOC,
      chromMAGMA_tf_MOC,
      chromMAGMA_tf_NMOC) %>%
  ungroup() %>%
  dplyr::select(MsigdB_NAME,
                GENE,
                GWAS_TYPE)

head(tf_list_final)

#write.table(x = tf_list_final, file = 'Data/nexus_tfs.6.15.21.txt', append = F, quote = T, sep = '\t', row.names = F, col.names = T)
```

